Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-08-01 writeup/hat-analysis.tex

For brevity, we provide an informal presentation of each guarantee
here (accompanied by appropriate references) but give a full set of
formal definitions in
 Appendix A.
% our extended Technical Report~\cite{hat-tr}.
In our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times but should
handle concurrent updates (i.e., overwrites) in accordance with the
total order for each item. \textit{PL-1} requires that writes to
different objects be ordered consistently across transactions,
prohibiting Adya's phenomenon $G0$ (also called ``Dirty
Writes''~\cite{ansicritique}). If we build a graph of transactions
with edges from one transaction to another and, when the former
overwrites the latter's write to the same object, then, under Read
Uncommitted, the graph should not contain cycles~\cite{adya}. Consider
the following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$. Read Uncommitted is easily
achieved via timestamping each update in a transaction with the same
timestamp (unique across transactions; e.g., combining a client's ID
with a sequence number) and applying a ``last writer wins'' conflict
reconciliation policy at each replica. Later properties will
strengthen Read Uncommitted.

@@

We have developed a more efficient RA algorithm, which we sketch here
and provide greater detail in Appendix B. We begin with our Read
Committed algorithm, but replicas wait to reveal new writes to readers
until all of the replicas for the final writes in the transaction have
received their respective writes (are \textit{pending
  stable}). Clients include additional metadata with each write: a
single timestamp for all writes in the transaction (e.g., as in Read
Uncommitted) and a list of items written to in the transaction. When a
client reads, the return value's timestamp and list of items form a
lower bound on the versions that the client should read for the other
items. When a client reads, it attaches a timestamp to its request
representing the current lower bound for that item. Replicas use this
timestamp to respond with either a write matching the timestamp or a
pending stable write with a higher timestamp. Servers keep two sets of
writes for each data item: the write with the highest timestamp that
is pending stable and a set of writes that are not yet pending
stable. This is entirely master-less and operations never block due to
replica coordination. Moreover, there are several possible
optimizations that, for brevity, we do not describe here.

@@

Consistent Read, Snapshot Isolation (including Parallel Snapshot
Isolation~\cite{walter}), and Cursor Stability guarantees are all
unavailable because they require preventing Lost Update phenomena.
Repeatable Read (defined by Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya~\cite{adya}) and One-Copy
Serializability~\cite{1sr} need to prevent both Lost Update and Write
Skew. Their phenomena prevention requirements mean that these
guarantees are inherently unachievable in a HAT system.

>>>>

For brevity, we provide an informal presentation of each guarantee
here (accompanied by appropriate references) but give a full set of
formal definitions in
\iftechreport
 Appendix A.
\else
 our extended Technical Report~\cite{hat-tr}.
\fi
In our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times but should
handle concurrent updates (i.e., overwrites) in accordance with the
total order for each item. \textit{PL-1} requires that writes to
different objects be ordered consistently across transactions,
prohibiting Adya's phenomenon $G0$ (also called ``Dirty
Writes''~\cite{ansicritique}). If we build a graph of transactions
with edges from one transaction to another and, when the former
overwrites the latter's write to the same object, then, under Read
Uncommitted, the graph should not contain cycles~\cite{adya}. Consider
the following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$. Read Uncommitted is easily
achieved by marking each of a transactions' writes with the same
timestamp (unique across transactions; e.g., combining a client's ID
with a sequence number) and applying a ``last writer wins'' conflict
reconciliation policy at each replica. Later properties will
strengthen Read Uncommitted.

@@

We have developed a more efficient RA algorithm, which we sketch here
and provide greater detail in
\iftechreport
 Appendix B.
\else
 our extended Technical Report~\cite{hat-tr}.
\fi
We begin with our Read
Committed algorithm, but replicas wait to reveal new writes to readers
until all of the replicas for the final writes in the transaction have
received their respective writes (are \textit{pending
  stable}). Clients include additional metadata with each write: a
single timestamp for all writes in the transaction (e.g., as in Read
Uncommitted) and a list of items written to in the transaction. When a
client reads, the return value's timestamp and list of items form a
lower bound on the versions that the client should read for the other
items. When a client reads, it attaches a timestamp to its request
representing the current lower bound for that item. Replicas use this
timestamp to respond with either a write matching the timestamp or a
pending stable write with a higher timestamp. Servers keep two sets of
writes for each data item: the write with the highest timestamp that
is pending stable and a set of writes that are not yet pending
stable. This is entirely master-less and operations never block due to
replica coordination. Moreover, there are several possible
optimizations that, for brevity, we do not describe here.

@@

Consistent Read, Snapshot Isolation (including Parallel Snapshot
Isolation~\cite{walter}), and Cursor Stability guarantees are all
unavailable because they require preventing Lost Update phenomena.
Repeatable Read (defined by Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya~\cite{adya}) and One-Copy
Serializability~\cite{1sr} need to prevent both Lost Update and Write
Skew. Their prevention requirements mean that these guarantees are
inherently unachievable in a HAT system.
